Towards a miniKanren with fair search strategies

### 2019-05-23 Paper/paper.tex
leading sentences

\usepackage{listings}
\lstset{
  language=Scheme,
  basicstyle=\ttfamily,
  morekeywords={run,conde,run*,defrel,==,fresh,disj,conj,disj+},
  alsodigit={!\$\%&*+-./:<=>?@^_~},
  morecomment=[l]{;,\#{}lang},
  mathescape=true,
  literate={-inf}{{$^\infty$}}{1} {/fair}{{$_{fair}$}}{3}
  		   {→}{{$\rightarrow$}}{1}
  		   {×}{{$\times$}}{1}
  		   {==}{{$\equiv$}}{1}
  		   {conde}{{\textcolor[rgb]{0,.3,.7}{\textbf{cond$^e$}}}}{4}
                   {appendo}{{append$^o$}}{6}
  		   {repeato}{{repeat$^o$}}{6}
  		   {g0}{{g$_0$}}{2}
  		   {g1}{{g$_1$}}{2}
  		   {g2}{{g$_2$}}{2}
  		   {gl}{{g$_l$}}{2}
  		   {gr}{{g$_r$}}{2}
  		   {DFSi }{{DFS$_i$}}{3}
  		   {DFSbi}{{DFS$_{bi}$}}{3}
  		   {DFSf }{{DFS$_f$}}{3}
  		   {BFSser}{{BFS$_{ser}$}}{3}
}
%% scheme-list .

@@

\newcommand{\DFSi }[0]{DFS$_\textrm{i}$}
\newcommand{\DFSf }[0]{DFS$_\textrm{f}$}
\newcommand{\DFSbi}[0]{DFS$_\textrm{bi}$}
\newcommand{\BFSser}[0]{BFS$_\textrm{ser}$}

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can
be fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.
The two non-standard depth-first searches are new. And we present a new, more 
efficient and simpler implementation of the standard breadth-first search. Using 
quantitative evaluation, we argue that the two new depth-first searches are 
competitive alternatives to the standard one, 
and that our standard breadth-first search implementation is more efficient than 
the current one.

@@

\NOTE[DPF]{Yes, that is what I mean. And furthermore, I think that changing opt
to imp (for improved) makes more sense.  We may be optimizing, but it seems more
like an improvement.  I did change some of them from optimized to improved.
In my letter wrt to using
BFSopt or BFSimp, it still makes more sense to me.  Perhaps it should be part
of the title that we have improved on the the standard BFS implementation.}
\NOTE[LKC]{Do you mean they should not come up in Abstract?}
\NOTE[DPF]{When you rewrite the abstract and drop the abbreviations, remember when
you are rewriting the Introduction and introducing the abbreviations, that you should
explain the ser of BFSser using a reference.}

@@

The miniKanren in TRS2 implements interleaving DFS (\DFSi), the cause of this 
unexpected result. With this search strategy, each clause takes half 
of its received computational resources and passes the other half to its 
following clauses, except for the last clause that takes all resources it 
receives. In the example above, the \texttt{a} clause takes half of all 
resourses. And the \texttt{b} clause takes a quarter. Thus \texttt{c} and 
\texttt{d} barely contribute to the result.

@@

\DFSi{} is sometimes powerful for an expert. By carefully organizing the order 
of \conde{} clauses, a miniKanren program can explore more ``interesting'' 
clauses than those uninteresting ones, and thus use computational resources 
efficiently. A little miniKanrener, however, may beg to differ--understanding
implementation details and fiddling with clause order is not the first
priority of a beginner.

There is another reason that miniKanren could use more search strategies than
just \DFSi. In many applications, there does not exist one order that serves all
purposes. For example, a relational dependent type checker contains
clauses for constructors that build data and clauses for eliminators that use
data. When the type checker is generating simple and shallow programs,
the clauses for constructors had better be at the top of the
\conde{} expression.
When performing proof searches for complicated programs, the clauses for 
eliminators had better be at the top of the \conde{} expression. With \DFSi, 
these two uses cannot be efficient at the same time. In fact, to make one use 
efficient, the other one must be more sluggish.

@@

To summarize our contributions, we
\begin{itemize}
	\item propose and implement balanced interleaving depth-first search 
	(\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement fair depth-first search (\DFSf{}), a 
	new search strategy with fair \disj{}.
	\item implement in a new way breath-first search (\BFSser), a search 
strategy with fair \disj{} and fair \conj{}. And we prove formally that our 
\BFSser{} implementation is semantically equivalent to \BFSser{}.
Our code runs faster in all benchmarks and we believe it is simpler.
\end{itemize}

@@

Now we elaborate fairness by running more queries about \repeato{}.

@@

\NOTE[DPF]{I think that should be said as directly as in your note.}
\NOTE[LKC]{We never use run* in this paper because fairness is only 
interesting when we have an unbounded number of answers. However, it is 
perfectly fine to use run* with any search strategies.}

@@

\BFSser{} puts states of the same cost in the same bag, and allocates
resources carefully among search spaces related to different bags such
that it produces answers in increasing order of cost. The \emph{cost}
of an answer is its depth in the search tree (i.e., the number of
calls to relations required to find them) \citep{seres1999algebra}. In
the following example, every answer is a list of a list of symbols,
where inner lists are the same. Here the cost of each answer is equal
to the length of its inner list plus the length of its outer list.

\NOTE[DPF]{Next to the last line of the above paragraph the word same
has too many meanings.  It could be have the same cost or it could be
failing to distinguish two list that would be the same if we substituted
a b for an a.  It needs to be clarified.  It might help to work out
the actual numbers}

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provides a comprehensive description of the 
miniKanren 
implementation but limited to unification constraints ($\equiv$).

\begin{figure}
	\lstinputlisting{Figures/conde.rkt}
	\caption{implementation of \conde}
	\label{conde}
\end{figure}

The definition of \conde{} (Fig.~\ref{conde}) is shared by all search 
strategies. It relates clauses disjunctively, and goals in the same 
clause conjunctively.

@@

\NOTE[DPF]{This came out fabulous.  I think I spotted some missed opportunities
later, which I will point out with additional notes.}
\NOTE[LKC]{I have turned those `subscripts' to real subscripts.}

Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: \texttt{s}s name 
states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name search spaces. Fig.~\ref{DFSi-0} shows the 
implementation of \disj. The 
first function, \texttt{disj2}, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \texttt{conj2}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The helper function 
\texttt{append-map$^\infty$} applies its input goal to states 
in its input search spaces and composes the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\texttt{disj2} (Fig.~\ref{fDFS}). The new \texttt{disj2} calls a new and 
fair version of \texttt{append$^\infty$}. \texttt{append$^\infty_{fair}$} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, set to 
\texttt{\#{}t}, which indicates that we haven't swapped
\texttt{s$^\infty$} and \texttt{t$^\infty$}. The swapping 
happens at 
the third \texttt{cond} clause in the helper, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \texttt{s$^\infty$} and 
\texttt{t$^\infty$} are ended with a thunk and we have swapped them. In this 
case, we construct a new thunk. The new thunk swaps two spaces back in the
recursive call. This is unnecessary for fairness. We do it to produce answers 
in a more natural order.

\NOTE[DPF]{When or where do you swap them back? This could be still clearer.}
\NOTE[LKC]{In the recursion}

@@

\BFSser{} has both fair \disj{} and fair \conj{}. Our implementation is based 
on \DFSf{} (not \DFSi{}). To implement \BFSser{} based on \DFSf{}, 
we need \texttt{append-map$^\infty_{fair}$} in addition to 
\texttt{append$^\infty_{fair}$}. 
The only difference between \texttt{append-map$^\infty$} and 
\texttt{append-map$^\infty_{fair}$} is that the latter calls 
\texttt{append$^\infty_{fair}$} instead of \texttt{append$^\infty$}.

The implementation is improvable in two ways. First, as mentioned in 
\autoref{sec:fairconj}, \BFSser{} puts answers in bags and answers of the same 
cost are in the 
same bag. In this implementation, however, it is unclear where this information 
is recorded. Second, \texttt{append$^\infty_{fair}$} is extravagant in memory 
usage. It 
makes $O(n+m)$ new \texttt{cons} cells every time, where $n$ and $m$ are the 
``length''s of input search spaces. We address these issues in the first 
subsection.

%Both \BFSser{} \citet{seres1999algebra} and \BFSopt{} produce answers in 
%increasing order of cost. So it is interesting to see if they are equivalent. 
%We prove so in Coq. The details are in the second subsection.

We prove formally our implementation is equivalent to the original one by 
\citet{seres1999algebra}. The details are in the second subsection.

@@

\begin{figure}
	\lstinputlisting{Figures/BFS-opt.rkt}	
	\caption{new and changed functions in optimized BFS that implements pure 
	features}
	\label{BFS-opt}
\end{figure}

As mentioned in \autoref{sec:fairconj}, \BFSser{} puts answers in bags and 
answers of the 
same cost are in the same bag. The cost
information is recorded subtly---the \texttt{car}s of a search space have cost 
0 (i.e. they are all in the same bag), and every thunk indicates an increment 
in cost. It is even more subtle that
\texttt{append$^\infty_{fair}$} and the \texttt{append-map$^\infty_{fair}$} 
respects the cost information. We make these facts more obvious by 
changing the type of search space, modifying related function definitions, 
and introducing a few more functions.

@@

We list functions related to the pure subset in Fig.~\ref{BFS-opt} (the 
others in Fig.~\ref{BFS-opt-cont}). We compare these functions with 
\citeauthor{seres1999algebra}'s implementation in our proof. The first three 
functions in Fig.~\ref{BFS-opt} are search space constructors. \texttt{none} 
makes an empty search space; \texttt{unit} makes a space from one answer; and 
\texttt{step} makes a space from a thunk. The remaining functions are as before.

@@

\citet{kiselyov2005backtracking} have demonstrated that a \emph{MonadPlus} 
hides in 
implementations of logic programming system. Our \BFSser{} implementation is 
not an exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, 
but takes arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, 
and \texttt{mplus}, respectively.

\begin{figure}
	\lstinputlisting{Figures/BFS-opt-cont.rkt}	
	\caption{new and changed functions in improved BFS that implement impure 
		features}
	\label{BFS-opt-cont}
\end{figure}

Functions implementing impure features are in Fig.~\ref{BFS-opt-cont}. The 
first function, \texttt{elim}, takes a space \texttt{s$^\infty$} and two 
continuations \texttt{ks} and \texttt{kf}. When \texttt{s$^\infty$} contains 
some 
answers, it calls \texttt{ks} with the first answer and the rest of the space. 
Otherwise, it calls \texttt{kf} with no argument. Here `s' means
`succeed' and `f' means `fail'. This function is similar to an eliminator of 
search spaces, hence the name. The remaining functions are as before.

\subsection{compare our implementation with the original one}

In this section, we compare the pure subset of our implementation with the 
original one. We focus on the pure subset because the original system is pure.
\citeauthor{seres1999algebra} represent search spaces with streams of 
lists of answers, where each list is a bag.

To compare efficiency, we translate \BFSser{}'s Haskell code into Racket (See supplements for the translated code). The translation is direct 
due to the similarity in both logic programming systems and search space 
representations. The translated code is longer and slower than our 
implementation. Details about differences in efficiency are in the table below.

We use Coq to show that the two implementations are 
semantically equivalent, (i.e., \texttt{(run n ? g)} 
produces the same result (See supplements for the formal proof).

@@

\begin{table}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline 
		benchmark & size & \DFSi{} & \DFSbi{} & \DFSf{} & 
		our \BFSser{} & original \BFSser{}  
		\\
		\hline
		\veryrecursiveo & 100000 &  579 &  793 & 2131 & 1438 & 3617 \\
		& 200000 & 1283 & 1610 & 3602 & 2803 & 4212 \\
		& 300000 & 2160 & 2836 &    - & 6137 &    - \\
		\hline 
		\appendo  & 100 &  31 &  41 &  42 &  31 &  68 \\ 
		& 200 & 224 & 222 & 221 & 226 & 218 \\ 
		& 300 & 617 & 634 & 593 & 631 & 622 \\ 
		\hline 
		\reverso & 10 &   5 &   3 &   3 &     38 &     85 \\ 
		& 20 & 107 &  98 &  51 &   4862 &   5844 \\
		& 30 & 446 & 442 & 485 & 123288 & 132159 \\ 
		\hline
		quine-1 & 1 &  71 &  44 & 69 & - & - \\ 
		& 2 & 127 & 142 & 95 & - & - \\ 
		& 3 & 114 & 114 & 93 & - & - \\ 
		\hline
		quine-2 & 1 & 147 & 112 &  56 & - & - \\ 
		& 2 & 161 & 123 & 101 & - & - \\ 
		& 3 & 289 & 189 & 104 & - & - \\ 
		\hline 
		'(I love you)-1 &  99 & 56 & 15 & 22 &  74 & 165 \\ 
		& 198 & 53 & 72 & 55 &  47 &  74 \\
		& 297 & 72 & 90 & 44 & 181 & 365 \\ 
		\hline
		'(I love you)-2 &  99 & 242 &  61 & 16 &  66 &  99 \\ 
		& 198 & 445 & 110 & 60 &  42 &  64 \\
		& 297 & 476 & 146 & 49 & 186 & 322 \\ 
		\hline 
	\end{tabular}
	\caption{The results of a quantitative evaluation: running times of 
	benchmarks 
		in milliseconds}
	\label{compare-efficiency}
\end{table}

Here we compare the efficiency of search strategies. A concise 
description is in Table~\ref{compare-efficiency}. A hyphen means ``running out of 
memory.'' The first two benchmarks are from 
TRS2. \reverso is from 
\citet{rozplokhas2018improving}. The next two benchmarks 
about quine are modified from a similar test case in \citet{byrd2017unified}. 
The modifications are made 
to circumvent the need for symbolic constraints (e.g. $\neq$, 
\texttt{absent$^o$}). Our version generates de 
Bruijnized expressions and prevents closures from being inside a list. The two 
benchmarks differ in the \conde{} clause order of their relational 
interpreters. 
The last two 
benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love 
you)}. This benchmark is from \citet{byrd2017unified}. Again, the 
sibling benchmarks differ in the \conde{} clause order of their relational 
interpreters. The first one 
has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the 
end, while the other has them at the beginning. We conjecture that \DFSi{} 
would 
perform badly in the second case because elimination rules complicate the 
problem when synthesizing (i.e., our evaluation supports our conjecture.)

>>>>

\usepackage{listings}
\lstset{
  language=Scheme,
  basicstyle=\ttfamily,
  morekeywords={run,conde,run*,defrel,==,fresh,disj,conj,disj+},
  alsodigit={!\$\%&*+-./:<=>?@^_~},
  morecomment=[l]{;,\#{}lang},
  mathescape=true,
  literate={-inf}{{$^\infty$}}{1} {/fair}{{$_{fair}$}}{3}
  		   {→}{{$\rightarrow$}}{1}
  		   {×}{{$\times$}}{1}
  		   {==}{{$\equiv$}}{1}
  		   {conde}{{\textcolor[rgb]{0,.3,.7}{\textbf{cond$^e$}}}}{4}
                   {appendo}{{append$^o$}}{6}
  		   {repeato}{{repeat$^o$}}{6}
  		   {g0}{{g$_0$}}{2}
  		   {g1}{{g$_1$}}{2}
  		   {g2}{{g$_2$}}{2}
  		   {gl}{{g$_l$}}{2}
  		   {gr}{{g$_r$}}{2}
  		   {DFSi }{{DFS$_\textrm{i}$}}{3}
  		   {DFSbi}{{DFS$_\textrm{bi}$}}{3}
  		   {DFSf }{{DFS$_\textrm{f}$}}{3}
}
%% scheme-list .

@@

\newcommand{\DFSi }[0]{DFS$_\textrm{i}$}
\newcommand{\DFSf }[0]{DFS$_\textrm{f}$}
\newcommand{\DFSbi}[0]{DFS$_\textrm{bi}$}
\newcommand{\BFS}[0]{BFS}
\newcommand{\BFSser}[0]{BFS$_\textrm{ser}$}
\newcommand{\BFSimp}[0]{BFS$_\textrm{imp}$}

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can
be fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.
The two non-standard depth-first searches are new. And we present a new, more 
efficient and simpler implementation of the standard breadth-first search. Using 
quantitative evaluation, we argue that the two new depth-first searches are 
competitive alternatives to the standard one, 
and that our breadth-first search implementation is more efficient than 
the current one.

@@

\NOTE[DPF]{... Perhaps it should be part of the title that we have improved on 
the the standard BFS implementation.}
\NOTE[DPF]{When you rewrite the abstract and drop the abbreviations, remember 
when you are rewriting the Introduction and introducing the abbreviations, that 
you should explain the ser of BFSser using a reference.}
\NOTE[LKC]{I have rewritten the abstract and explain the `ser' in the ending 
paragraph of introduction, where we list contributions.}

@@

The miniKanren in TRS2 implements interleaving DFS (\DFSi), the cause of this 
unexpected result. With this search strategy, each \conde{} clause takes half 
of its received computational resources and passes the other half to its 
following clauses, except for the last clause that takes all resources it 
receives. In the example above, the \texttt{a} clause takes half of all 
resourses. And the \texttt{b} clause takes a quarter. Thus \texttt{c} and 
\texttt{d} barely contribute to the result.

@@

\DFSi{} is sometimes powerful for an expert. By carefully organizing the order 
of \conde{} clauses, a miniKanren program can explore more ``interesting'' 
clauses than those uninteresting ones, and thus use computational resources 
efficiently.

% A little miniKanrener, however, may beg to differ--understanding
% implementation details and fiddling with clause order is not the first
% priority of a beginner.

\DFSi{} is not always the best choice. For instance, it might be less 
desirable for little miniKanreners -- understanding implementation details and 
fiddling with clause order is not their first priority. 
There is another reason that miniKanren could use more search strategies than
just \DFSi. In many applications, there does not exist one order that serves all
purposes. For example, a relational dependent type checker contains
clauses for constructors that build data and clauses for eliminators that use
data. When the type checker is generating simple and shallow programs,
the clauses for constructors had better be at the top of the
\conde{} expression.
When performing proof searches for complicated programs, the clauses for 
eliminators had better be at the top of the \conde{} expression. With \DFSi, 
these two uses cannot be efficient at the same time. In fact, to make one use 
efficient, the other one must be more sluggish.

@@

To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
a new search strategy with fair \disj{}.
	\item implement in a new way the standard breath-first search (\BFS), a 
search strategy with fair \disj{} and fair \conj{}. We name the current 
implementation by \citeauthor{seres1999algebra} as \BFSser{}, and our 
new implementation as \BFSimp{}. We prove formally that they are semantically 
equivalent. \BFSimp{}, however, runs faster in all benchmarks and is shorter, 
thus being an \textbf{imp}rovement. 
\end{itemize}

@@

\NOTE[LKC]{The following paragraph is extended to include an explanation on 
\texttt{run*}}

Now we elaborate fairness by running more queries about \repeato{}. We never use 
\texttt{run*} in this paper because fairness is more interesting when we have 
an unbounded number of answers. However, it is perfectly fine to use 
\texttt{run*} with any search strategies.

@@

\NOTE[DPF]{I think that should be said as directly as in your note.}
\NOTE[LKC]{I state this fact at the beginning of this section.}
\NOTE[LKC]{We never use run* in this paper because fairness is only 
interesting when we have an unbounded number of answers. However, it is 
perfectly fine to use run* with any search strategies.}

@@

\BFS{} puts states of the same cost in the same bag, and allocates
resources carefully among search spaces related to different bags such
that it produces answers in increasing order of cost. The \emph{cost}
of an answer is its depth in the search tree (i.e., the number of
calls to relations required to find them) \citep{seres1999algebra}. In
the following example, every answer is a list of a list of symbols,
where inner lists in the same outer list are identical. Here the cost of each 
answer is equal
to the length of its inner list plus the length of its outer list. For example,
the cost of \texttt{((a) (a))} is $1 + 2 = 3$.

\NOTE[DPF]{Next to the last line of the above paragraph the word same
has too many meanings.  It could be have the same cost or it could be
failing to distinguish two list that would be the same if we substituted
a b for an a.  It needs to be clarified.  It might help to work out
the actual numbers}
\NOTE[LKC]{I have rewritten the relevant sentences.}

@@

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provides a comprehensive description of the 
miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: \texttt{s}s name 
states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name search spaces. Fig.~\ref{DFSi-0} shows the 
implementation of \disj. The 
first function, \texttt{disj2}, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \texttt{conj2}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The helper function 
\texttt{append-map$^\infty$} applies its input goal to states 
in its input search spaces and composes the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say \conj{} is also right-associative.
% \begin{figure}
% 	\lstinputlisting{Figures/conde.rkt}
% 	\caption{implementation of \conde}
% 	\label{conde}
% \end{figure}
% 
% The definition of \conde{} (Fig.~\ref{conde}) is shared by all search 
% strategies. It relates clauses disjunctively, and goals in the same 
% clause conjunctively.

\NOTE[DPF]{This came out fabulous. I think I spotted some missed opportunities
later, which I will point out with additional notes.}
\NOTE[LKC]{I think we can simply remove \conde{} code and its accompany 
description.}

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\texttt{disj2} (Fig.~\ref{fDFS}). The new \texttt{disj2} calls a new and 
fair version of \texttt{append$^\infty$}. \texttt{append$^\infty_{fair}$} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, set to 
\texttt{\#{}t}, which indicates that we haven't swapped
\texttt{s$^\infty$} and \texttt{t$^\infty$}. The swapping 
happens at 
the third \texttt{cond} clause in the helper, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \texttt{s$^\infty$} and 
\texttt{t$^\infty$} are ended with a thunk, and that we have swapped them. In 
this case, we construct a new thunk. The new thunk swaps two spaces back in the
recursive call to \texttt{loop}. This is unnecessary for fairness. We do it to 
produce answers in a more natural order.

\NOTE[DPF]{When or where do you swap them back? This could be still clearer.}
\NOTE[LKC]{Improved.}

@@

\BFS{} has both fair \disj{} and fair \conj{}. An easy implementation is based 
on \DFSf{} (not \DFSi{}): (1) replace \texttt{append$^\infty$} with 
\texttt{append$^\infty_{fair}$} in \texttt{append-map$^\infty$}'s body (2) 
rename \texttt{append-map$^\infty$} to \texttt{append-map$^\infty_{fair}$}.

% Our implementation is based 
% on \DFSf{} (not \DFSi{}). To implement \BFS{} based on \DFSf{}, 
% we need \texttt{append-map$^\infty_{fair}$} in addition to 
% \texttt{append$^\infty_{fair}$}. 
% The only difference between \texttt{append-map$^\infty$} and 
% \texttt{append-map$^\infty_{fair}$} is that the latter calls 
% \texttt{append$^\infty_{fair}$} instead of \texttt{append$^\infty$}.

\NOTE[LKC]{Applying the above changes won't result in \BFSser, but yet 
another implementation of \BFS.}

% This implementation is improvable in two ways. First, as mentioned in 
% \autoref{sec:fairconj}, \BFS{} puts answers in bags and answers of the same 
% cost are in the same bag. In this implementation, however, it is unclear where 
% this information is recorded. Second, \texttt{append$^\infty_{fair}$} is 
% extravagant in memory usage. It makes $O(n+m)$ new \texttt{cons} cells every 
% time, where $n$ and $m$ are the ``length''s of input search spaces.

This implementation is improvable in two ways. 
First, as mentioned in \autoref{sec:fairconj}, \BFS{} puts answers in bags and 
answers of the same cost are in the same bag. In this implementation, however, 
we manage cost information subtly---the \texttt{car}s of a search space have 
cost 0 (i.e. they are all in the same bag), and every thunk indicates an 
increment in cost. It is even more subtle that \texttt{append$^\infty_{fair}$} 
and the \texttt{append-map$^\infty_{fair}$} respects the cost information. 
Second, \texttt{append$^\infty_{fair}$} is extravagant in memory usage. It 
makes $O(n+m)$ new \texttt{cons} cells every time, where $n$ and $m$ are the 
``length''s of input search spaces.

\NOTE[LKC]{Should I point out that \DFSf{} also has the miserable space 
extravagance?}

In the following subsections, we firstly describe \BFSimp{}, an improved \BFS{} 
implementation which manages cost information in a more clear way and is less 
extravagant in memory usage. Then we compare \BFSimp{} with \BFSser{}, the 
current implementation by \citeauthor{seres1999algebra}.

@@

\begin{figure}
	\lstinputlisting{Figures/BFSimp.rkt}	
	\caption{new and changed functions in optimized BFS that implements pure 
	features}
	\label{BFSimp}
\end{figure}

% As mentioned in \autoref{sec:fairconj}, \BFS{} puts answers in bags and 
% answers of the same cost are in the same bag. 
% In the easy implementation, cost information is recorded subtly---the 
% \texttt{car}s of a search space have cost 0 (i.e. they are all in the same 
% bag),
% and every thunk indicates an increment in cost. It is even more subtle that
% \texttt{append$^\infty_{fair}$} and the \texttt{append-map$^\infty_{fair}$} 
% respects the cost information. We make these facts more obvious by changing 
% the type of search space, modifying related function definitions, and 
% introducing a few more functions.

We make the cost information more clear by changing the type of search space, 
modifying related function definitions, and introducing a few more functions.

@@

We list functions related to the pure subset in Fig.~\ref{BFSimp}. The first 
three functions are search space constructors. \texttt{none} makes an empty 
search space; \texttt{unit} makes a space from one answer; and \texttt{step} 
makes a space from a thunk. The remaining functions are as before. We compare 
these functions with \BFSser{} in our proof.

@@

\NOTE[LKC]{I do notice that in the below paragraph the leading sentence is not 
at the beginning (but at the second position). Putting the current first 
sentence at the beginning, however, looks more fun to me.}

\citet{kiselyov2005backtracking} have demonstrated that a \emph{MonadPlus} 
hides in implementations of logic programming system. \BFSimp{} is not an 
exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, but takes 
arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, and 
\texttt{mplus}, respectively.

\begin{figure}
	\lstinputlisting{Figures/BFSimp-cont.rkt}	
	\caption{new and changed functions in improved BFS that implement impure 
		features}
	\label{BFSimp-cont}
\end{figure}

Functions implementing impure features are in Fig.~\ref{BFSimp-cont}. The 
first function, \texttt{elim}, takes a space \texttt{s$^\infty$} and two 
continuations \texttt{kf} and \texttt{ks}. When \texttt{s$^\infty$} contains 
no answers, it calls \texttt{kf} with no argument. Otherwise, it calls 
\texttt{ks} with the first answer and the rest of the space. Here  `f' means 
`fail' and `s' means `succeed'. This function is similar to an eliminator of 
search spaces, hence the name. The remaining functions are as before.

\NOTE[LKC]{I swap \texttt{ks} and \texttt{kf} to mimic \texttt{ind-List}}

\subsection{compare \BFSimp{} with \BFSser{}}

\NOTE[LKC]{I have rewritten this subsection.}

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We proof in Coq that these two search strategies 
are are semantically equivalent, (i.e., \texttt{(run n ? g)} produces the same 
result (See supplements for the formal proof). To compare efficiency, we 
translate \BFSser{}'s Haskell code into Racket (See supplements for the 
translated code). The translation is direct due to the similarity in two 
relational programming systems. The translated code is longer than \BFSimp{}. 
And it runs slower in all benchmarks. Details about differences in efficiency 
are in \autoref{compare-efficiency}.

@@

\begin{table}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline 
		benchmark & size & \DFSi{} & \DFSbi{} & \DFSf{} & 
		\BFSimp{} & \BFSser{}  
		\\
		\hline
		\veryrecursiveo & 100000 &  579 &  793 & 2131 & 1438 & 3617 \\
		& 200000 & 1283 & 1610 & 3602 & 2803 & 4212 \\
		& 300000 & 2160 & 2836 &    - & 6137 &    - \\
		\hline 
		\appendo  & 100 &  31 &  41 &  42 &  31 &  68 \\ 
		& 200 & 224 & 222 & 221 & 226 & 218 \\ 
		& 300 & 617 & 634 & 593 & 631 & 622 \\ 
		\hline 
		\reverso & 10 &   5 &   3 &   3 &     38 &     85 \\ 
		& 20 & 107 &  98 &  51 &   4862 &   5844 \\
		& 30 & 446 & 442 & 485 & 123288 & 132159 \\ 
		\hline
		quine-1 & 1 &  71 &  44 & 69 & - & - \\ 
		& 2 & 127 & 142 & 95 & - & - \\ 
		& 3 & 114 & 114 & 93 & - & - \\ 
		\hline
		quine-2 & 1 & 147 & 112 &  56 & - & - \\ 
		& 2 & 161 & 123 & 101 & - & - \\ 
		& 3 & 289 & 189 & 104 & - & - \\ 
		\hline 
		'(I love you)-1 &  99 & 56 & 15 & 22 &  74 & 165 \\ 
		& 198 & 53 & 72 & 55 &  47 &  74 \\
		& 297 & 72 & 90 & 44 & 181 & 365 \\ 
		\hline
		'(I love you)-2 &  99 & 242 &  61 & 16 &  66 &  99 \\ 
		& 198 & 445 & 110 & 60 &  42 &  64 \\
		& 297 & 476 & 146 & 49 & 186 & 322 \\ 
		\hline 
	\end{tabular}
	\caption{The results of a quantitative evaluation: running times of 
	benchmarks 
		in milliseconds}
	\label{compare-efficiency}
\end{table}

In this section, we compare the efficiency of search strategies. A concise 
description is in Table~\ref{compare-efficiency}. A hyphen means ``running out
of memory.'' The first two benchmarks are from 
TRS2. \reverso is from 
\citet{rozplokhas2018improving}. The next two benchmarks 
about quine are modified from a similar test case in \citet{byrd2017unified}. 
The modifications are made 
to circumvent the need for symbolic constraints (e.g. $\neq$, 
\texttt{absent$^o$}). Our version generates de 
Bruijnized expressions and prevents closures from being inside a list. The two 
benchmarks differ in the \conde{} clause order of their relational 
interpreters. 
The last two 
benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love 
you)}. This benchmark is from \citet{byrd2017unified}. Again, the 
sibling benchmarks differ in the \conde{} clause order of their relational 
interpreters. The first one 
has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the 
end, while the other has them at the beginning. We conjecture that \DFSi{} 
would 
perform badly in the second case because elimination rules complicate the 
problem when synthesizing (i.e., our evaluation supports our conjecture.)

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{} is just 
as efficient in all benchmarks but \veryrecursiveo. Both implementations of BFS 
have obvious overhead in many cases. Among the three variants of DFS, which all 
all have unfair \conj{}), \DFSf{} is most resistant to clause permutation, 
followed by \DFSbi{} then \DFSi{}. Among the two implementation of BFS, our 
improved \BFSimp{} constantly performs as well or better. Interestingly, every 
strategy with fair \disj{} suffers in \veryrecursiveo{} and \DFSf{} performs 
well elsewhere. Therefore, this benchmark might be a special case. Fair \conj{} 
imposes considerable overhead constantly except in \appendo. The reason might be 
that strategies with fair \conj{} tend to keep more intermediate answers in the 
memory.
